\LARGE \bf Preparation of Papers for IEEE Sponsored Conferences \& Symposia*

### 2017-06-08 graph_abs.tex
stuff
section 3 done

Let $\inputsignal$ = $(t_0, u_0)$, $\ldots$, $(t_N,u_N)$ be a
timed-sequence of inputs, where $t_0 = 0$, and for all $i > 0$, $t_i -
t_{i-1} = \Delta$, and $u_i \in \Inputs$.  We can extend the forward
simulation map to a sequence of states to describe the application of
an input signal to excite the system as follows:
$\simmap_\Delta(\initstate, \mathbf{u} = \contState_N$, if for all $i  \in
[0,N-1]$, $\simmap_\Delta(\contState_i, u_i,t_i) = \contState_{i+1}$.

% An alternative view of the forward simulation-map is as a
% parameterized reachability relation $\reach_\Delta{t,u} \subseteq
% \ContStates \times \ContStates$ defined as below:
% \[
% (\contState,\contState') \in \reach_\Delta{t,u}\text{ iff } \contState' = \simmap_\Delta(\contState,t,u).
% \]

@@

% \mypara{System Assumptions:} 
% We require that $\System$ can always be
% simulated forward in time with deterministic results. This in turn
% requires existence and uniqueness of trajectories over a finite time
% horizon $[0,T]$, which can be guaranteed by Lipschitz continuity of
% the vector field in each underlying hybrid mode, ruling away issues
% such as finite escape times~\cite{Meiss/2007/Differential}, and having
% well-defined semantics for simulations at switching boundaries. While
% these assumptions hold for the dynamical system examples which we
% consider in this paper, such assumptions can be restrictive for
% general control system models. When analyzing a system that does not
% obey theoretical conditions that allow deterministic simulation, we
% assume that the underlying tool used for numerical approximations of
% system behaviors uses deterministic simulation semantics.  Finally, we
% assume full observability of the system state, validate our results
% only against the $\mathsc{sim}^\System$ function (as the analytic closed-form
% representation of the system dynamics is assumed to be unavailable).
% From here on, we drop the subscript $\System$ whenever it is clear
% from the context. Also, to ease the presentation, we conflate the
% inputs with the state and denote them by $\vx$.
% 
\subsection{Graph Abstraction}

We observe that the parameterized reachability relation described in
the previous section induces an existential discrete abstraction of
the system similar to the one considered in \cite{zutshi2014multiple}.
We briefly review this discrete abstraction and show how falsification
of systems reduces to a path-search in the resulting transition
system. We make the observation that this discrete abstraction is
actually a Kripke structure, \ie, a transition system defined as the
tuple $(\Cells, \CellEdges, \CellLabelingFunction)$, where $\Cells$ is
a finite set of cells, $\CellEdges \subseteq \Cells \times \Cells$ is
an unlabeled transition relation, and $\CellLabelingFunction$ is a
function that assigns one or more labels to the cells from a finite
set of labels.

In what follows, we review an $\epsilon$-precision transition system
abstraction of a system $\System$ (denoted as
$\TSAbstraction{\epsilon}{\System}$) as follows. We define a
quantization function $\quant_\epsilon$ that maps a state $\contState$ in
$\ContStates$ to $10^{-\epsilon}*\lfloor \contState*10^\epsilon \rfloor$.
This function induces an equivalence relation $\contState_i
\equiv_\epsilon \contState_j$ iff $\quant_\epsilon(\contState_i) =
\quant_\epsilon(\contState_j)$. Now consider the quotient structure
$\ContStates/\equiv_\epsilon$. We treat each equivalence class in the
quotient structure as a cell. We discover the transitions in the
abstraction using simulations. Given a numerical simulator for the
model with a forward simulation map $\simmap_\Delta$, we add an edge
from cell $C$ to cell $C'$ iff:
\begin{itemize}
\item
$\simmap_\Delta(\contState, u, t)$ = $\contState'$, 
\item
$\quant_\epsilon(\contState)$ = $C$, and,
\item
$\quant_\epsilon(\contState') = C'$.
\end{itemize}

While our framework is general enough to reason about arbitrary
temporal logic properties of trajectories, we restrict our attention
to safety properties, where we assume that there is a specified set of
states $\UnsafeStates \subseteq \ContStates$ that are unsafe.  Thus,
our label set has only two elements:
$\setof{\init,\unsafe}$. We then define $\CellLabelingFunction$ as
follows: 
\begin{equation*}
\CellLabelingFunction(C) = \left\{
\begin{array}{l@{\hspace{1em}}l}
\init   & \exists \contState \in \InitStates\text{ s.t.  }\quant_\epsilon(\contState) = C  \\
\unsafe & \exists \contState \in \UnsafeStates\text{ s.t.  }\quant_\epsilon(\contState) = C  \\
\{\}    & \text{otherwise.}    
\end{array}\right.
\end{equation*}

@@

We remark that our abstraction is {\em neither an overapproximation
nor an underapproximation} of the continuous-time continuous-space
reachability relation between states. The purpose of this abstraction
is to discover falsifying trajectories; thus, this technique explores
the abstraction to get abstract falsifying trajectories starting at an
initial abstract state that lead to error, and to confirm that such an
abstract trajectory is not spurious, it performs a concretization step
by performing an actual simulation of the system. If the abstract
counterexample trajectories are found to be spurious, then
$\TSAbstraction{\epsilon}{\System}$ can be refined, basically by
increasing $\epsilon$. We summarize this procedure in
Algorithm~\ref{algo:cegar}.

The biggest drawback of Algorithm~\ref{algo:cegar} is that with
increasing $\epsilon$, the size of $\Cells$ in
$\TSAbstraction{\epsilon}{\System}$ increases exponentially.  Thus,
every subsequent run of the falsification algorithm is slower. In the
next section, we propose an alternative abstraction that enriches the
basic abstraction proposed here by adding labels to the transitions.
We can then perform refinement by enriching the transition relation,
rather than increasing $\epsilon$.

>>>>

% Let $\inputsignal$ = $(t_0, u_0)$, $\ldots$, $(t_N,u_N)$ be a
% timed-sequence of inputs, where $t_0 = 0$, and for all $i > 0$, $t_i -
% t_{i-1} = \Delta$, and $u_i \in \Inputs$.  We can extend the forward
% simulation map to a sequence of states to describe the application of
% an input signal to excite the system as follows:
% $\simmap_\Delta(\initstate, \mathbf{u}) = \contState_N$, if for all $i
% \in [0,N-1]$, $\simmap_\Delta(\contState_i, u_i,t_i) =
% \contState_{i+1}$.
% 
% An alternative view of the forward simulation-map is as a
% parameterized reachability relation $\reach_\Delta{t,u} \subseteq
% \ContStates \times \ContStates$ defined as below:
% \[
% (\contState,\contState') \in \reach_\Delta{t,u}\text{ iff } \contState' = \simmap_\Delta(\contState,t,u).
% \]

@@

% \mypara{System Assumptions:} 
% We require that $\System$ can always be
% simulated forward in time with deterministic results. This in turn
% requires existence and uniqueness of trajectories over a finite time
% horizon $[0,T]$, which can be guaranteed by Lipschitz continuity of
% the vector field in each underlying hybrid mode, ruling away issues
% such as finite escape times~\cite{Meiss/2007/Differential}, and having
% well-defined semantics for simulations at switching boundaries. While
% these assumptions hold for the dynamical system examples which we
% consider in this paper, such assumptions can be restrictive for
% general control system models. When analyzing a system that does not
% obey theoretical conditions that allow deterministic simulation, we
% assume that the underlying tool used for numerical approximations of
% system behaviors uses deterministic simulation semantics.  Finally, we
% assume full observability of the system state, validate our results
% only against the $\mathsc{sim}^\System$ function (as the analytic closed-form
% representation of the system dynamics is assumed to be unavailable).
% From here on, we drop the subscript $\System$ whenever it is clear
% from the context. Also, to ease the presentation, we conflate the
% inputs with the state and denote them by $\vx$.
% 
\subsection{$\epsilon$-precision Transition System Abstraction}

We now show how we can abstract a hybrid dynamical system $\System$
which has trajectories in dense-time by a discrete transition system.
A discrete transition system $\mathcal{T}$ can be defined as the tuple
$(\Cells, \CellEdges, \Cells_\init, \CellLabelingFunction)$, where
$\Cells$ is a finite set of cells, $\CellEdges \subseteq \Cells \times
\Cells$ is an unlabeled transition relation, and
$\CellLabelingFunction$ is a function that assigns one or more labels
to the cells from a finite set $L$ of labels. The operational
semantics of the transition system are as follows: the system starts
in some cell in $\Cells_\init$. At each step, if the transition system
is in cell $C$, it transitions to cell $C'$ if $(C,C') \in
\CellEdges$. A finite-time trace property $\varphi$ of the system is
some subset of the free monoid $(2^L)^*$. We say that $\mathcal{T}
\not\models \varphi$ if there exists some trace of $\mathcal{T}$ such
that the sequence of labels in that trace does not belong to
$\varphi$.

We now give an example of a specific kind of transition system
abstraction that we can build given $\System$ and the simulation map
$\simmap_\Delta$.  We call the transition system an
$\epsilon$-precision transition system and denote it by
$\TSAbstraction{\epsilon}{\System}$.  Let $\epsilon  = 10^{-k}$, for
some positive integer $k$.  We define a quantization function
$\quant_\epsilon$ that maps a state $\contState$ in $\ContStates$ to
$10^{-k}*\lfloor \contState*10^k \rfloor$. This transformation
essentially truncates the decimal representation of the numeric value
of the state $\state$ to at most $k$ digits after the decimal point.
This function induces an equivalence relation $\contState_i
\equiv_\epsilon \contState_j$ iff $\quant_\epsilon(\contState_i) =
\quant_\epsilon(\contState_j)$.

Using the equivalence relation $\equiv_\epsilon$, we can now define
the tuple $(\Cells, \CellEdges, \CellLabelingFunction)$ for
$\TSAbstraction{\epsilon}{\System}$ by considering the quotient
structure $\ContStates/\equiv_\epsilon$.  Essentially, each
equivalence class in the quotient structure is a cell in $\Cells$.
For cells $C$ and $C'$ in $\TSAbstraction{\epsilon}{\System}$, the
edge $(C,C')$ can be an element of $\CellEdges$ iff:
$\simmap_\Delta(\contState, u, t)$ = $\contState'$,
$\quant_\epsilon(\contState)$ = $C$, and,
$\quant_\epsilon(\contState') = C'$.

We assume that there is a specified set of states $\UnsafeStates
\subseteq \ContStates$ that are unsafe\footnote{Our framework is
general enough to reason about richer class of correctness properties
for trajectories such as bounded-time LTL properties. In this case, we
assume the LTL property is defined over Boolean-valued predicates over
$\ContStates$. The labeling function $\CellLabelingFunction$ would
then label cells with the predicates that evaluate to $\mathit{true}$
for some state in the cell.}, and thus restrict our attention to
simple safety properties. Thus, our label set has only two elements:
$\setof{\init,\unsafe}$. We then define $\CellLabelingFunction$ as
$\CellLabelingFunction(C) = \init$ if $C$ contains some initial state,
$\CellLabelingFunction(C) = \unsafe$ if $C$ contains some unsafe
state, and empty otherwise.

In what follows, we review the procedure
$\mathsf{ScatterAndSimulate}$($\System$,$\simmap_\Delta$,$\epsilon$)
from \cite{zutshi2014multiple} that, given the $\System$, the forward
simulation map $\simmap_\Delta$, and a precision $\epsilon$ creates an
$\epsilon$-precision transition system abstraction of $\System$.  The
essence of the procedure is a breadth-first exploration of
$\ContStates$ to build $\TSAbstraction{\epsilon}{\System}$ =
$(\Cells,\CellEdges,\CellLabelingFunction)$. It consists of the
following steps:
\begin{enumerate}
\item
Add a set of random samples from the initial set of states to a queue
(\ie some states $\contState$ at time $0$).
\item
Pop the head of the queue (say state $\contState$ at time $t$), identify
the corresponding cell by applying $\quant_\epsilon$, and add it to
$\Cells$. 
\item
Obtain $\contState' = \simmap_\Delta(\contState, t, u)$. Obtain a a
fixed number of states in the neighborhood of $\contState'$, and add
$\contState'$ and these states (all time $t+\Delta$) to the queue. 
\item
Add an edge between the cells corresponding to $\contState$ and 
$\contState'$.
\item
Go to step 2 till either the queue is empty or a fixed number of cells
or edges is added to $\TSAbstraction{\epsilon}{\System}$.
\end{enumerate}

@@

We remark that our abstraction is {\em neither an overapproximation
nor an underapproximation} of the continuous-time continuous-space
reachability relation between states. The purpose of the
$\epsilon$-precision transition system is to discover falsifying
trajectories. This can be accomplished by
Algorithm~\ref{algo:s3camBMC} that uses CounterExample-Guided
Abstraction and Refinement (CEGAR).

Here, we construct $\TSAbstraction{\epsilon}{\System}$
(Line~\ref{algoline:construct}), we check if there is a path in
$\TSAbstraction{\epsilon}{\System}$ that leads to a cell labeled
unsafe. This can be done by a bounded model checking procedure
(Line~\ref{algoline:bmc}) that explores at most $\ell$ transitions of
$\TSAbstraction{\epsilon}{\System}$ (\ie, traces that are at most
$\ell\Delta$ seconds in length). The model checker may return a set of
counterexample trajectories $\Pi$. If this set is empty
(Line~\ref{algoline:nocex}), then either the exploration performed by
$\mathsf{ScatterAndSimulate}$ is not enough or the system could be
safe. In the former case, the user can increase the scope of
exploration for $\mathsf{ScatterAndSimulate}$.  The next step is to
check if the abstract trajectories are spurious, by performing a
concretization step by performing an actual simulation of the system
(Line~\ref{algoline:checkspurious}). If all abstract counterexample
trajectories are found to be spurious, then
$\TSAbstraction{\epsilon}{\System}$ can be refined, basically by
increasing $\epsilon$.
\begin{algorithm}[t] 
\DontPrintSemicolon
\caption{CEGAR for $\TSAbstraction{\epsilon}{\System}$\label{algo:s3camBMC}}
\KwIn{$\System$, $\simmap_\Delta$, $\epsilon_0$ , Smallest Precision $\epsilonmin$} 
$\epsilon \assign \epsilon_0$ \;
\While{$\epsilon \ge \epsilonmin$}{
    $\TSAbstraction{\epsilon}{\System}$ $\assign$ 
    $\mathsf{ScatterAndSimulate}$($\System$,$\simmap_\Delta$,$\epsilon$)
    \nllabel{algoline:construct} \;
    $\Pi$ $\assign$ $\mathsf{BoundedModelCheck}$($\TSAbstraction{\epsilon}{\System}$) 
    \nllabel{algoline:bmc} \;
    \lIf {$\Pi = \emptyset$}{ return FAIL \nllabel{algoline:nocex} } 
    \lIf {$\mathsf{checkIfConcretizable}(\Pi)$}{ return VIOLATION \nllabel{algoline:checkspurious} }
    $\epsilon$ $\assign$ $\frac{\epsilon}{10}$ \nllabel{algoline:refine} \;
}
\end{algorithm}

The biggest drawback of Algorithm~\ref{algo:s3camBMC} is that with
increasing $\epsilon$, the size of $\Cells$ in
$\TSAbstraction{\epsilon}{\System}$ increases exponentially.  Thus,
every subsequent run of the falsification algorithm is slower. In the
next section, we propose an alternative abstraction that enriches the
basic abstraction proposed here by adding labels to the transitions.
We can then perform refinement by enriching the transition relation,
rather than increasing $\epsilon$.
